Flexible Runtime Code Specialization for the Kernel

### 2013-11-10 VEE-14/paper.tex
Working on the DBT section

The rest of the paper describes Granary in more detail. \Cref{sec:modes} describes how to efficiently implement mixed-mode execution. \Cref{sec:policies} describes a new method for dynamically changing how code is instrumented. \Cref{sec:reify} describes how to integrate static analysis information into a DBT system. Finally, \Cref{sec:eval} evaluates Granary's performance.

@@

Granary translates and instruments kernel binaries one basic block at a time. In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction. Entrypoint basic blocks (e.g. interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time as execution ``discovers'' them. Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}.

@@

Basic blocks in Granary's code cache contain x86-64 binary instructions and meta-data describing those instructions. The meta-data that Granary records in each basic block includes: \begin{inparaenum}[i)]
	\item the length in bytes of the original and translated basic blocks;
	\item the address of the first instruction in the original basic block; and
	\item the policy information (e.g., $P_{\mathit{call\_entry}}$, $P_{\mathit{after\_entry}}$) used to instrument the basic block.
\end{inparaenum} This meta-data can be queried and extended by Granary instrumentation tools. For example, Granary's CFG-building tool represents each node in the inter- and intra-procedural CFGs as an extension of the meta-data stored for each basic block. The meta-data is also queried by interrupt handlers when deciding how to handle interrupts in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.

@@

An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}. Granary includes configurable levels of transparency to address the trade-off between transparency and overhead. By default, Granary uses a relaxed transparency model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility. By default, Granary exposes the following three artifacts: code cache addresses as function and interrupt return addresses, and wrapped module function addresses in shared data structures.

\vspace{-3pt}\paragraph{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary defaults to inlining \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses. Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks. If Granary is configured to use transparent return addresses, then implicit attaching and detaching are disabled because they depend directly on returning to instrumented module or native kernel code. Granary will fall back to slow attaching to handle kernel-to-module returns, and indirect branch lookups for module-to-module returns.

\vspace{-3pt}\paragraph{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts. A special case arises for page fault exceptions that occur within specific kernel functions that access user space data. Granary handles this special case in a similar way to the kernel. Thus, maintaining return address transparency with respect to interrupts in instrumented module code has no value.

\vspace{-3pt}\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions. This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behaviour under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.

\section{Policy-Driven Instrumentation}\label{sec:policies}

@@

\section{Conclusion}\label{sec:conclusion}
We created Granary to address the challenges of instrumenting binary kernel modules.  Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.  A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.  Granary also supports policy-driven instrumentation, which allows tools to specialize their instrumentation based on the context in which code executes.  Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level. 
Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.

>>>>

The rest of the paper describes Granary in more detail. \Cref{sec:dbt} describes how Granary is implemented.

% \Cref{sec:modes} describes how to efficiently implement mixed-mode execution. \Cref{sec:policies} describes a new method for dynamically changing how code is instrumented. \Cref{sec:reify} describes how to integrate static analysis information into a DBT system. Finally, \Cref{sec:eval} evaluates Granary's performance.

@@

Granary translates and instruments kernel binaries one basic block at a time. In Granary, a basic block is a sequence of instructions ending in a conditional branch, \texttt{ret}, or \texttt{jmp}, but not a \texttt{call} instruction. Entrypoint basic blocks (e.g. interrupt vectors, system call entrypoints, module initializers) are translated ahead-of-time, and all other basic blocks are translated just-in-time (JIT) as execution ``discovers'' them. Translated basic blocks are linked together and stored in a globally accessible \emph{code cache}.

@@

Basic blocks in Granary's code cache contain x86-64 binary instructions. Associated with each basic block is meta-data describing the instructions of the basic block, and the instrumentation policy used to instrument those instructions. Basic block meta-data includes: \begin{inparaenum}[i)]
	\item the length in bytes of the original and translated basic blocks;
	\item the address of the first instruction in the original basic block; and
	\item the instrumentation policy.
\end{inparaenum} Basic block meta-data is queried and extended by Granary instrumentation tools. For example, Granary's control-flow graph build tool, \texttt{cfg}, extends the basic block meta-data to track per-block execution counts. Block meta-data is also queried by interrupt handlers when deciding how to handle interrupts and exceptions in instrumented code, and by debuggers (e.g., \texttt{gdb}) to give contextual information about a translated basic block.

An instrumentation system is transparent if, given the same inputs, the instrumented and uninstrumented versions of a program behave in the same way \cite{Transparency}. Granary includes configurable levels of transparency to address the trade-off between transparency and overhead. By default, Granary uses a relaxed transparency model for efficiency, flexibility and increased visibility, but if instrumentation for some module requires transparency then it can be enabled at the cost of increased overheads and decreased visibility. By default, Granary exposes the following two artifacts: code cache addresses as function and interrupt return addresses.

\paragraph{Function return addresses}\label{para:return_address_transparency} Unlike most DBT frameworks, Granary inlines \texttt{call} instructions into basic blocks, which exposes code cache addresses to instrumented module code in the form of return addresses. Inlining \texttt{call} instructions avoids return-address mispredictions and unnecessary control-flow by building longer basic blocks.

\paragraph{Interrupt return addresses} Kernel interrupts are not sensitive to return addresses and hence relaxing return address transparency has no effect on interrupts. A special case arises for page fault exceptions that occur within specific kernel functions that access user space data \cite{btkernel}. The Linux kernel records ranges of code addresses that are permitted to fault in an ``exception table'' data structure, and on a page fault exception, checks if the interrupt return address belongs to one of the pre-defined ranges. Granary records any exception table entries as part of each basic block's meta-data, so that a page fault in a basic block can be mapped to the correct native faulting program counter.

%\paragraph{Module function wrappers} Granary does not maintain transparency when its type wrappers replace pointers to module functions with pointers to wrapped module functions. This is perhaps the riskiest break in transparency within Granary because a module might treat a function pointer as being representative of the module being in some particular state.\footnote{We have not yet encountered a module where changing pointers to module functions into module wrappers altered the module's behaviour under instrumentation.} If Granary is configured to use transparent module function pointer addresses in shared data structures then fast attaching is disabled and tools will not have access to the same level of static program information.

\subsection{Re-entrancy}

\paragraph{Code cache}
Granary's code cache and JIT translation mechanism is fully re-entrant. Context-switches into Granary disable interrupts and switch execution onto a CPU-private stack. Context switches out of Granary and into the code cache or native code returns executuin to the native stack and re-enables interrupts. JIT translation, however, presents a problem: context-switches save or restore the entire register state, so as not to disturb the execution of the instrumented program. This raises the question: when resolving the target of a direct or indirect control-flow instruction, how should execution be directed to the correct location?

Granary's solution to this problem is two-fold. Like \texttt{btkernel}, Granary generates ``edge'' code for each direct control-flow instruction (e.g. \texttt{call}, \texttt{jmp}, \texttt{jcc}).

\paragraph{Instrumentation}

@@

\section{Context-Driven Instrumentation}\label{sec:policies}

@@

\section{Conclusion}\label{sec:conclusion}
%We created Granary to address the challenges of instrumenting binary kernel modules.  Granary provides mixed-mode execution to remove the overheads of DBT for uninstrumented kernel code.  A relaxed transparency model further improves performance and allows greater visibility into the interactions between module and kernel code.  Granary also supports policy-driven instrumentation, which allows tools to specialize their instrumentation based on the context in which code executes.  Finally, Granary exposes high-level static analysis information to dynamic instrumentation tools, making it easy to match low-level memory reads with specific fields in data structures at the source code level.

%Together, Granary's features simplify the development of powerful kernel module analysis tools, while delivering lower overheads than previous kernel dynamic binary instrumentation solutions.
